---
title: "rpick Notes"
output:
  html_notebook: default
---

```{r}
source("https://raw.githubusercontent.com/nimids/R-Projects/main/sourceGitHub.R")
```

# Random Pick:

Picks n items from an urn filled with choices, with or without replacement.

- [Usage](#usage)
- [Arguments](#arguments)
- [Examples](#examples)
- [Plots](#plots)

## Usage:
```{r, eval=FALSE}
rpick(
  n = 1,
  choices = c("A","B","C"),
  replace = TRUE,
  verbose = FALSE
)
```

## Arguments:
--------  --------------------------------------------------
n         number of draws
choices   vector of choices
replace   if TRUE, choices are put back before the next draw
verbose   if TRUE, forces rpick to print out it's self
--------  --------------------------------------------------																	

## Examples:

Do one sample:
```{r}
n <- 10
choices <- c(1,1,1,2,2,3)
pickResult <- rpick(n, choices)
```

Print out the result:
```{r}
print(pickResult)
```

Look at raw values:
```{r}
value(pickResult)
```

Do lots of one samples:
```{r}
samples <- 1000

samplesTable <- do(samples) * rpick(n, choices)

print(samplesTable)
```


  We don't need choices, replace and result, also do adds some extra columns we
don't need, .row and .index we could just ignore them. or we can use mutate to
set the columns to NULL removing them.
```{r}
samplesTable %>%
  mutate(
    choices = NULL,
    replace = NULL,
    result = NULL,
    .row = NULL,
    .index = NULL
  ) ->
  samplesTable
print(samplesTable)
```
  
Print out counts of each result:
```{r}
{
  cat("count numOf1:")
  print(table(samplesTable$numOf1))
  cat("count numOf2:")
  print(table(samplesTable$numOf2))
  cat("count numOf3:")
  print(table(samplesTable$numOf3))
}
```

## Plots:
### Count Bar Plot - Num of 1s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_counts(~ numOf1) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, n, length = n+1), limits = c(-0.5, n+0.5)))
)
```

### Count Bar Plot - Num of 2s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_counts(~ numOf2) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, n, length = n+1), limits = c(-0.5, n+0.5)))
)
```

### Count Bar Plot - Num of 3s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_counts(~ numOf3) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, n, length = n+1), limits = c(-0.5, n+0.5)))
)
```

### Percents Bar Blot - Num of 3s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_percents(~ numOf3) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, n, length = n+1), limits = c(-0.5, n+0.5)))
)
```

### Proportion Bar Blot - Num of 3s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_props(~ numOf3) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, n, length = n+1), limits = c(-0.5, n+0.5)))
)
```

### Proportion Bar Blot - Proportions of 3s:
```{r fig.height=6.75, fig.width=12}
print(
  samplesTable %>%
    gf_props(~ numOf3/n) %>%
    gf_refine(scale_x_continuous(breaks = seq(0, 1, length = n+1), limits = c(-0.5, n+0.5)/n))
)
```

  If we wanted something like the number of 1s + number of 2s we can use mutate
to build a new table.

```{r}
samplesTable %>%
  mutate(
    numOf1Plus2 = numOf1 + numOf2,
    propOf1Plus2 = (numOf1 + numOf2)/n
  )
  
```


  If we save that table we can use it to plot later or if we only plan on using
it once we can just send it right to gf_props, gf_counts or gf_percents

```{r fig.height=6.75, fig.width=12}
samplesTable %>%
  mutate(
    propOf1Plus2 = (numOf1 + numOf2)/n
  ) %>%
  gf_props(~ propOf1Plus2) %>%
  gf_refine(scale_x_continuous(breaks = seq(0, 1, length = n+1), limits = c(-0.5, n+0.5)/n))
```
  Or we can use the formula we give gf_props to do simple operations like
adding, subtracting, multiplying and dividing.

```{r fig.height=6.75, fig.width=12}
samplesTable %>%
  gf_props(~ 1 - (numOf1 + numOf2)/n) %>%
  gf_refine(scale_x_continuous(breaks = seq(0, 1, length = n+1), limits = c(-0.5, n+0.5)/n))
```
